---
title: "Part A – SQL Analysis"
---

------------------------------------------------------------------------

# Introduction

This first part demonstrates the use of SQL to query a SQLite database with three different core SQL queries. The results are displayed using formatted and paginated tables.

------------------------------------------------------------------------

```{r}
## Load the necessary libraries

library(DBI)
library(RSQLite)
library(dplyr)
library(DT)

## Connect to the SQLite database
con <- dbConnect(
  SQLite(),
  "virtual_production.db"
)
## Check tables exist and we are connected to the DB

dbListTables(con)
```

# SQL Queries

## 1st core query

List the total budget allocated for productions in each country along with the count of projects per country sorted by total budget (descending)

```{r}
query_budget_country <- "
SELECT 
  c.Country AS Country,
  COUNT(p.ProductionID) AS Number_of_Productions,
  SUM(p.Budget) AS Total_Budget
FROM Productions p
JOIN Clients c
  ON p.ClientID = c.ClientID
GROUP BY c.Country
ORDER BY Total_Budget DESC
"

budget_country_df <- dbGetQuery(con, query_budget_country)

## Display Results as a Paginated Table

datatable(
  budget_country_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

### **Explanation 1st query**

This query joins the `Productions` and `Clients` tables to aggregate production budgets by country. The total budget and number of productions are calculated using `SUM` and `COUNT`, and the results are sorted in descending order of total budget to highlight countries with the highest investment in virtual production.

## 2nd core query

Next is the query that lists the average development time for productions, categorized by the number of assets

```{r}
query_avg_dev_time <- "
SELECT
  AssetCount,
  AVG(DevelopmentDays) AS Avg_Development_Days
FROM (
  SELECT
    p.ProductionID,
    JULIANDAY(p.EndDate) - JULIANDAY(p.StartDate) AS DevelopmentDays,
    COUNT(a.AssetID) AS AssetCount
  FROM Productions p
  LEFT JOIN Environments e
    ON p.ProductionID = e.ProductionID
  LEFT JOIN Assets a
    ON e.EnvironmentID = a.EnvironmentID
  GROUP BY p.ProductionID
)
GROUP BY AssetCount
ORDER BY AssetCount
"

avg_dev_time_df <- dbGetQuery(con, query_avg_dev_time)

## Display as a Paginated Table


datatable(
  avg_dev_time_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

### **Explanation 2nd query**

This query calculates the development duration for each production by subtracting the start date from the end date using SQLite’s `JULIANDAY` function. A subquery is used to count the number of assets associated with each production. The outer query then groups productions by asset count and computes the average development time, highlighting how asset volume relates to production duration.

## 3rd core query

Final 3rd query lists the top three crew members based on the number of successful productions they’ve been involved in.

```{r}
query_top_crew <- "
SELECT
  cm.Name AS Crew_Member,
  COUNT(p.ProductionID) AS Successful_Productions
FROM CrewMembers cm
JOIN ProductionCrew pc
  ON cm.CrewID = pc.CrewID
JOIN Productions p
  ON pc.ProductionID = p.ProductionID
WHERE p.Status IN ('Post-production', 'Completed')
GROUP BY cm.CrewID, cm.Name
ORDER BY Successful_Productions DESC
LIMIT 3
"

top_crew_df <- dbGetQuery(con, query_top_crew)

## Display the Results

datatable(
  top_crew_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)

```

### **Explanation 3rd query**

This query identifies the top three crew members based on their involvement in successful productions. Success is defined as productions with a status of Post-production or Completed. The query joins crew members to productions via the production–crew linking table, aggregates the number of successful productions per crew member, and orders the results in descending order to highlight the most experienced contributors.

# SQL Concepts

## 1st concept

Filter using `LIKE` combine conditions with `OR`

### Example scenario

List crew members whose role contains *Unreal* **or** *Camera*

```{r}
query_like_or <- "
SELECT
  Name,
  Role,
  ExperienceYears
FROM CrewMembers
WHERE Role LIKE '%Unreal%'
   OR Role LIKE '%Camera%'
"

like_or_df <- dbGetQuery(con, query_like_or)

### Display the Table

datatable(
  like_or_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

### **Explanation 1st concept**

This query demonstrates the use of the `LIKE` operator combined with the `OR` condition to filter textual data. Crew members are selected if their role contains either the term *Unreal* or *Camera*, allowing flexible pattern matching across multiple criteria.

## 2nd concept

Filter using `SELECT DISTINCT` with `ORDER BY`

### Example scenario

List distinct production genres in alphabetical order

```{r}
query_distinct_order <- "
SELECT DISTINCT
  Genre
FROM Productions
ORDER BY Genre ASC
"

distinct_genres_df <- dbGetQuery(con, query_distinct_order)

### Display Table

datatable(
  distinct_genres_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

### **Explanation 2nd concept**

This query uses \`SELECT DISTINCT\` to return unique production genres, eliminating duplicate values. The \`ORDER BY\` clause sorts the results alphabetically, improving readability and supporting categorical analysis.

## 3rd concept

Subquery with `SELECT`

### Example scenario

List productions whose budget is above the **average budget**

```{r}
query_subquery <- "
SELECT
  Title,
  Budget
FROM Productions
WHERE Budget > (
  SELECT AVG(Budget)
  FROM Productions
)
ORDER BY Budget DESC
"

above_avg_budget_df <- dbGetQuery(con, query_subquery)


### Display Table

datatable(
  above_avg_budget_df,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

### **Explanation 3rd concept**

This query demonstrates the use of a subquery to calculate the average production budget. The outer query then filters productions whose budget exceeds this average, illustrating how subqueries can be used to perform comparative analysis within SQL.
